Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2022-01-15 response-to-reviews.Rmd

**The units in the table are in %nm i.e. % wavelength**

Paragraphs 3.1 and 3.2:
I thought that the comparison between intra species and interspecies gradients is very interesting. I wondered if there are more studies to expand it. One key article that I know is (Lloyd et al., 2010) which showed that within tree gradients and between tree gradients are similar for a lot of leaf traits in tropical forests. I think this is one of the reasons why DGCMs are relatively accurate even if the canopy description is quite simple. (The vertical gradients in photosynthetic parameters are scaled on Na gradients, which are the same intra and between species in this study).

L429 I am not fully sure if I understood this sentence, consider maybe developing the idea. Do you mean that the variation is higher in canopy made of multi species than on the canopy of single species? If so, I didn’t understand the comparison with understory species.

**We removed this sentence because yes, it is confusing, and the points addressed from comments above provide substantial information for inter- and intra-canopy leaf traits.**

@@

*yes, I think this sentence is confusing now that I am reading it again, need to work on this*

L470 A thought: Stomatal conductance is highly correlated to photosynthetic capacity (Wong et al., 1979) so the gradients in conductance are expected to follow the gradients in photosynthetic capacity. The water use efficiency (gs/A) or the slope parameter of conductance models (Medlyn et al., 2011) could change vertically inside the profiles. I don’t know if there are a lot of studies on the vertical variation of the water use efficiency or the slope parameter, but that could be something interesting to mention. See for example (Buckley, 2021).

I also think you could describe a little bit more the hydraulic constraints that increase with height. See for example (Koch et al., 2004). I think there are also other papers discussing the effect of height on water potential and stomatal conductance. You mention some aspects of it later, around L 636.

*need to add more water water use efficiency*

**We added the following sentences related to hydraulic contraints with height:** 
**Further, tree height also plays a role in determining the gradient of $g_s$ vertically, which has implications for $T_leaf$. With an increase in tree height, water travels a longer distance to meet the evaporative demands in sunlit canopy leaves. Longer water path length, high irradiance and VPD imposes hydraulic constraints in sunlit canopy leaves (Schäfer et al., 2000; Ambrose et al., 2010). Therefore, although typically $g_s$ increases with light, upper canopy leaves of taller trees have lower $g_s$ and leaf water potential to enable high leaf water-use efficiency (Rijkers et al., 2000; Kenzo et al., 2015). This, added to the tendency for sun leaves to have higher $T_leaf$, implies that high $T_air$ should decrease $g_s$ of canopy leaves more than understory leaves, particularly when water availability is limited.**

### 2022-01-15 vertical-thermal-review.Rmd
minor edits

Further, tree height also plays a role in determining the gradient of $g_s$ vertically, which has implications for $T_{leaf}$. With an increase in tree height, water travels a longer distance to meet the evaporative demands in sunlit canopy leaves. Longer water path length, higher irradiance and VPD imposes hydraulic constraints in sunlit canopy leaves [@schaferEffectTreeHeight2000; @ambroseEffectsHeightTreetop2010]. Therefore, although typically $g_s$ increases with light, upper canopy leaves of taller trees have lower $g_s$ and leaf water potential to enable high leaf water-use efficiency [@kenzoHeightrelatedChangesLeaf2015; @rijkersEffectTreeHeight2000a]. This, added to the tendency for sun leaves to have higher $T_{leaf}$, implies that high $T_{air}$ should decrease $g_s$ of canopy leaves more than understory leaves, particularly when water availability is limited.

>>>>

Further, tree height also plays a role in determining the gradient of $g_s$ vertically, which has implications for $T_{leaf}$. With an increase in tree height, water travels a longer distance to meet the evaporative demands in sunlit canopy leaves. Longer water path length, higher irradiance and VPD imposes hydraulic constraints in sunlit canopy leaves [@schaferEffectTreeHeight2000; @ambroseEffectsHeightTreetop2010]. Therefore, although typically $g_s$ increases with light, due to hydraulic constraints, upper canopy leaves of taller trees have lower $g_s$ and leaf water potential to enable high leaf water-use efficiency [@kenzoHeightrelatedChangesLeaf2015; @rijkersEffectTreeHeight2000a]. This, added to the tendency for sun leaves to have higher $T_{leaf}$, implies that high $T_{air}$ should decrease $g_s$ of canopy leaves more than understory leaves, particularly when water availability is limited.
